\thetitle

### 2020-11-09 sun_phases/spin_qudit_tomo.tex
removed one minor comment and added another

Computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
This complexity of this task can be greatly reduced by the fact that the degree $\ell$ of the transition operator $T_{\ell,0}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell,m} \ket{\ell m} \obra{T_{\ell m}},
\end{align}
block-diagonalizes the measurement matrix:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$ blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for additional details on efficiently computing $\S_{V\ell}^2$ [\red{todo: write tutorial}]).
Altogether, if we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure [\red{REF}]).
[\red{todo: add figure showing runtimes as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].
For comparison, we note that existing algorithms for signal recovery on the sphere \cite{mcewen2011novel, rauhut2011sparse, alem2012sparse, khalid2014optimaldimensionality} similarly have $O(d^3)$ parallel runtime with $\abs{V}\gtrsim d$.

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by experimental sources of noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} and derive more fundamental error bounds in Appendix \ref{sec:error}.
Therein, we find that if measurement error is dominated by shot noise, then the estimate $\tilde\rho_V$ built from $n$ independent measurements of spin projection along every axis $\v v\in V$ approximates the true state $\rho$ to within a mean squared distance $\epsilon_V^2/n$, i.e.
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2} < \f{\epsilon_V^2}{n},
  \label{eq:bound}
\end{align}
where the {\it error scale} $\epsilon_V^2$ is
\begin{align}
  \epsilon_V^2 \equiv
  \f1d \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2,
  \label{eq:scale}
\end{align}
and $\gamma_\ell$ is defined using Clebsch-Gordan coefficients:
\begin{align}
  \gamma_\ell
  \equiv \f{\max_\mu\tau_{\ell\mu} - \min_\mu\tau_{\ell\mu}}{2},
  &&
  \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu}.
\end{align}
For comparison, we note that the error scale $\epsilon_V^2<\S_V^2/2$.

>>>>

Computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} requires constructing the measurement matrix $M_V$ and finding its singular values.
This complexity of this task can be greatly reduced by the fact that the degree $\ell$ of the transition operator $T_{\ell,0}$ is preserved under rotations, which implies that the unitary
\begin{align}
  U \equiv \sum_{\ell,m} \ket{\ell m} \obra{T_{\ell m}},
\end{align}
block-diagonalizes the measurement matrix:
\begin{align}
  U M_V = \sum_\ell \op{\ell} \otimes M_{V\ell},
\end{align}
where the $\p{2\ell+1}\times\abs{V}$ blocks are
\begin{align}
  M_{V\ell} \equiv \sum_{m,\v v} \ket{m} \obk{T_{\ell m}|T_{\v v\ell,0}} \bra{\v v}
  = \sum_{m,\v v} D^\ell_{m,0}\p{\v v} \op{m}{\v v}.
\end{align}
As the singular values of $M_V$ are invariant under unitary transformations, it follows that
\begin{align}
  \S_V^2 = \sum_\ell \S_{V\ell}^2,
  &&
  \S_{V\ell}^2 \equiv \norm{M_{V\ell}^{-1}}^2,
\end{align}
where $M_{V\ell}^{-1}$ is the right inverse of $M_{V\ell}$.
Constructing the block $M_{V\ell}$ and computing its singular values takes at most $O(\abs{V}d^2)$ time (see Appendix \ref{sec:compute} for additional details on efficiently computing $\S_{V\ell}^2$ [\red{todo: write tutorial}]).
Altogether, if we assume that $\abs{V}\sim d$, then computing $\S_V^2$ to certify the bound in Eq.~\eqref{eq:bound_eps} takes $O(d^4)$ serial or $O(d^3)$ parallel runtime (see Figure [\red{REF}]).
[\red{todo: add figure showing runtimes as a function of $d$ with $\abs{V}=2d-1$; modify text appropriately after making the figure}].

The assumption that observables can be estimated up to uncorrelated noise with maximal variance $\epsilon^2$, summarized by Eq.~\eqref{eq:error_assumption}, is reasonable when measurement error is dominated by experimental sources of noise.
However, this assumption breaks down when measurement error is limited by fundamental quantum (shot) noise.
We relax the assumption of Eq.~\eqref{eq:error_assumption} and derive more fundamental error bounds in Appendix \ref{sec:error}.
Therein, we find that if measurement error is dominated by shot noise, then the estimate $\tilde\rho_V$ built from $n$ independent measurements of spin projection along every axis $\v v\in V$ approximates the true state $\rho$ to within a mean squared distance $\epsilon_V^2/n$, i.e.
\begin{align}
  \Bbk{\norm{\tilde\rho_V-\rho}^2} < \f{\epsilon_V^2}{n},
  \label{eq:bound}
\end{align}
where the {\it error scale} $\epsilon_V^2$ is
\begin{align}
  \epsilon_V^2 \equiv
  \f1d \sum_{\ell>0} \p{2\ell+1} \gamma_\ell^2 \S_{V\ell}^2,
  \label{eq:scale}
\end{align}
and $\gamma_\ell$ is defined using Clebsch-Gordan coefficients:
\begin{align}
  \gamma_\ell
  \equiv \f{\max_\mu\tau_{\ell\mu} - \min_\mu\tau_{\ell\mu}}{2},
  &&
  \tau_{\ell\mu} \equiv \bk{s\mu;\ell,0|s\mu}.
\end{align}
If $d$ even or $\ell$ is odd, then $\gamma_\ell=\max_\mu\tau_{\ell\mu}$.
For comparison, we note that the error scale $\epsilon_V^2<\S_V^2/2$.
